DATA 621 Homework #4

### 2019-11-13 hw4/HW_4_Corey.Rmd
Renamed file

---
title: "DATA 621 Homework #4"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
require(gridExtra)
library(Amelia)
library(kableExtra)
library(caret)
library(DMwR)
library(scales)
library(RColorBrewer)
library(ROCR)
# Thank you Stack Overflow!
# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})
knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

```{r}
df <- read.csv("./data/insurance_training_data.csv")
evaluation <- read.csv("./data/insurance-evaluation-data.csv")

strip_dollars <- function(x){
  x <- as.character(x)
  x <- gsub(",", "", x)
  x <- gsub("\\$", "", x)
  as.numeric(x)
}
```

@@

```{r}
strip_dollars <- function(x){
  x <- as.character(x)
  x <- gsub(",", "", x)
  x <- gsub("\\$", "", x)
  as.numeric(x)
}

fix_data_types <- function(messy_df){
  messy_df %>%
    rowwise() %>%
    mutate(INCOME = strip_dollars(INCOME),
           HOME_VAL = strip_dollars(HOME_VAL),
           BLUEBOOK = strip_dollars(BLUEBOOK),
           OLDCLAIM = strip_dollars(OLDCLAIM)) %>%
    ungroup()
}

na_bad_values <- function(messy_df){
  messy_df %>%
    rowwise() %>%
    mutate(CAR_AGE = ifelse(CAR_AGE < 0, NA, CAR_AGE))%>%
    ungroup()
}

df$TARGET_FLAG <- factor(df$TARGET_FLAG)

df <- df %>%
  fix_data_types() %>%
  na_bad_values()
evaluation <- evaluation %>%
  fix_data_types() %>%
  na_bad_values()
```

@@

```{r}
# Function to add features

create_features <- function(d){
  d %>%
    mutate(LOG_INCOME = log(INCOME + 1),
           LOG_HOME_VAL = log(HOME_VAL + 1),
           AVG_CLAIM = ifelse(CLM_FREQ > 0, OLDCLAIM / CLM_FREQ, 0),
           PRIOR_ACCIDENT = factor(ifelse(OLDCLAIM == 0 & AVG_CLAIM == 0, 0, 1)),
           COLLEGE_EDUCATED = factor(ifelse(EDUCATION %in% c("Bachelors", "Masters", "PhD"), 1, 0)),
           URBAN_DRIVER = factor(ifelse(URBANICITY == "Highly Urban/ Urban", 1, 0)),
           YOUNG_MALE = factor(ifelse(SEX == "M" & AGE < 25, 1, 0)),
           YOUNG = factor(ifelse(AGE < 25, 1, 0)),
           RED_SPORTS_CAR = factor(ifelse(CAR_TYPE == "Sports Car" & RED_CAR == "yes", 1, 0)),
           HAS_KIDS = factor(ifelse(HOMEKIDS == 0, 0, 1)),
           KID_DRIVERS = factor(ifelse(KIDSDRIV == 0, 0, 1)))
}

df <- create_features(df)
evaluation <- create_features(evaluation)
```

@@

```{r, echo=FALSE}
over_sample_train %>% 
  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
  ggplot(aes(x=TARGET_FLAG, fill=TARGET_FLAG)) +
  geom_bar() + scale_y_continuous() + scale_fill_brewer(palette="Set1") +
  theme_light() +
  theme(legend.position = "none") +
  labs(x="TARGET_FLAG", y="# Observations")

over_sample_train$TARGET_FLAG <- factor(over_sample_train$TARGET_FLAG)
```

@@

```{r, null_prefix = TRUE, comment=NA}
baseline_model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT, family = binomial(link = "logit"), over_sample_train)
summary(baseline_model)

results <- evaluate_model(baseline_model, test)
```

@@

```{r, null_prefix = TRUE, comment=NA}
risk_taker_model <- glm(TARGET_FLAG ~ RED_SPORTS_CAR + YOUNG_MALE, family = binomial(link = "logit"), over_sample_train)
summary(risk_taker_model)

risk_taker_results <- evaluate_model(risk_taker_model, test)
```

@@

```{r}
traditional_model <- glm(TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX +
                           REVOKED + MVR_PTS + TRAVTIME + CAR_USE,
             family = binomial(link = "logit"), over_sample_train)

summary(traditional_model)

model_results <- evaluate_model(traditional_model, test)
```

@@

```{r}
# Get complete cases
cases <- train %>%
  select(YOUNG,MSTATUS,PRIOR_ACCIDENT,SEX,REVOKED,MVR_PTS, TRAVTIME, CAR_USE) %>%
  complete.cases()

temp <- train[cases,]

# Use 5-fold cross-validation
train_control <- trainControl(method="cv",number=5, sampling="up")

traditional_cv <- train(form = TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX +
                           REVOKED + MVR_PTS + TRAVTIME + CAR_USE,
                        method = "glm", family = "binomial", 
                        data = temp, trControl = train_control)

traditional_cv
```

```{r}
# Evaluating the model
eval <- data.frame(actual = test$TARGET_FLAG, predicted = predict(traditional_cv,newdata=test,type="raw"), prob = 
predict(traditional_cv,newdata=test,type="prob"))

confusionMatrix(eval$predicted, eval$actual, positive = "1")

pred <- prediction(eval$prob.1, eval$actual)
auc <- performance(pred, measure = "auc")@y.values[[1]]
perf <- performance(pred,"tpr","fpr")
plot(perf,main="ROC Curve", sub=paste0("AUC: ",round(auc,3)))
```

@@

```{r, null_prefix = TRUE, comment=NA}
#"KIDSDRIV"         "AGE"             
#[6] "HOMEKIDS"         "YOJ"              "INCOME"           "PARENT1"          "HOME_VAL"        
#[11] "MSTATUS"          "SEX"              "EDUCATION"        "JOB"              "TRAVTIME"        
#[16] "CAR_USE"          "BLUEBOOK"         "TIF"              "CAR_TYPE"         "RED_CAR"         
#[21] "OLDCLAIM"         "CLM_FREQ"         "REVOKED"          "MVR_PTS"          "CAR_AGE"         
#[26] "URBANICITY"       "LOG_INCOME"       "LOG_HOME_VAL"     "AVG_CLAIM"        "PRIOR_ACCIDENT"  
#[31] "COLLEGE_EDUCATED" "URBAN_DRIVER"     "YOUNG_MALE"       "RED_SPORTS_CAR"   "HAS_KIDS"        
#[36] "KID_DRIVERS"     
model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT + KID_DRIVERS + MSTATUS + INCOME +
             SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER,
             family = binomial(link = "logit"), over_sample_train)
summary(model)

model_results <- evaluate_model(model, test)
```

>>>>

---
title: "DATA 621 Homework #4"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    toc_depth: 3
    code_folding: "hide"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
require(gridExtra)
library(Amelia)
library(kableExtra)
library(caret)
library(DMwR)
library(scales)
library(RColorBrewer)
library(ROCR)
# Thank you Stack Overflow!
# A Prefix nulling hook.
# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")
# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {
  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }
  default_output_hook( x, options )
})
knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
```

```{r}
df <- read.csv("./data/insurance_training_data.csv")
evaluation <- read.csv("./data/insurance-evaluation-data.csv")
strip_dollars <- function(x){
  x <- as.character(x)
  x <- gsub(",", "", x)
  x <- gsub("\\$", "", x)
  as.numeric(x)
}
```

@@

```{r}
strip_dollars <- function(x){
  x <- as.character(x)
  x <- gsub(",", "", x)
  x <- gsub("\\$", "", x)
  as.numeric(x)
}
fix_data_types <- function(messy_df){
  messy_df %>%
    rowwise() %>%
    mutate(INCOME = strip_dollars(INCOME),
           HOME_VAL = strip_dollars(HOME_VAL),
           BLUEBOOK = strip_dollars(BLUEBOOK),
           OLDCLAIM = strip_dollars(OLDCLAIM)) %>%
    ungroup()
}
na_bad_values <- function(messy_df){
  messy_df %>%
    rowwise() %>%
    mutate(CAR_AGE = ifelse(CAR_AGE < 0, NA, CAR_AGE))%>%
    ungroup()
}
df$TARGET_FLAG <- factor(df$TARGET_FLAG)
df <- df %>%
  fix_data_types() %>%
  na_bad_values()
evaluation <- evaluation %>%
  fix_data_types() %>%
  na_bad_values()
```

@@

```{r}
# Function to add features
create_features <- function(d){
  d %>%
    mutate(LOG_INCOME = log(INCOME + 1),
           LOG_HOME_VAL = log(HOME_VAL + 1),
           AVG_CLAIM = ifelse(CLM_FREQ > 0, OLDCLAIM / CLM_FREQ, 0),
           PRIOR_ACCIDENT = factor(ifelse(OLDCLAIM == 0 & AVG_CLAIM == 0, 0, 1)),
           COLLEGE_EDUCATED = factor(ifelse(EDUCATION %in% c("Bachelors", "Masters", "PhD"), 1, 0)),
           URBAN_DRIVER = factor(ifelse(URBANICITY == "Highly Urban/ Urban", 1, 0)),
           YOUNG_MALE = factor(ifelse(SEX == "M" & AGE < 25, 1, 0)),
           YOUNG = factor(ifelse(AGE < 25, 1, 0)),
           RED_SPORTS_CAR = factor(ifelse(CAR_TYPE == "Sports Car" & RED_CAR == "yes", 1, 0)),
           HAS_KIDS = factor(ifelse(HOMEKIDS == 0, 0, 1)),
           KID_DRIVERS = factor(ifelse(KIDSDRIV == 0, 0, 1)))
}
df <- create_features(df)
evaluation <- create_features(evaluation)

df$years <- ifelse(df$TIF >5,5,df$TIF)

df$ClaimFreqPeryear <- df$CLM_FREQ/df$years

df$ClaimAMTPerYear <- df$OLDCLAIM/df$years

df$years <- NULL
```

@@

```{r, echo=FALSE}
over_sample_train %>% 
  mutate(TARGET_FLAG = as.factor(TARGET_FLAG)) %>%
  ggplot(aes(x=TARGET_FLAG, fill=TARGET_FLAG)) +
  geom_bar() + scale_y_continuous() + scale_fill_brewer(palette="Set1") +
  theme_light() +
  theme(legend.position = "none") +
  labs(x="TARGET_FLAG", y="# Observations")
over_sample_train$TARGET_FLAG <- factor(over_sample_train$TARGET_FLAG)
```

@@

```{r, null_prefix = TRUE, comment=NA}
baseline_model <- glm(TARGET_FLAG ~ PRIOR_ACCIDENT, family = binomial(link = "logit"), over_sample_train)
summary(baseline_model)
results <- evaluate_model(baseline_model, test)
```

@@

```{r, null_prefix = TRUE, comment=NA}
risk_taker_model <- glm(TARGET_FLAG ~ RED_SPORTS_CAR + YOUNG_MALE, family = binomial(link = "logit"), over_sample_train)
summary(risk_taker_model)
risk_taker_results <- evaluate_model(risk_taker_model, test)
```

@@

```{r}
traditional_model <- glm(TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX +
                           REVOKED + MVR_PTS + TRAVTIME + CAR_USE,
             family = binomial(link = "logit"), over_sample_train)
summary(traditional_model)
model_results <- evaluate_model(traditional_model, test)
```

@@

```{r}
# Get complete cases
cases <- train %>%
  select(YOUNG,MSTATUS,PRIOR_ACCIDENT,SEX,REVOKED,MVR_PTS, TRAVTIME, CAR_USE) %>%
  complete.cases()
temp <- train[cases,]
# Use 5-fold cross-validation
train_control <- trainControl(method="cv",number=5, sampling="up")
traditional_cv <- train(form = TARGET_FLAG ~ YOUNG + MSTATUS + PRIOR_ACCIDENT + SEX +
                           REVOKED + MVR_PTS + TRAVTIME + CAR_USE,
                        method = "glm", family = "binomial", 
                        data = temp, trControl = train_control)
traditional_cv
```

```{r}
# Evaluating the model
eval <- data.frame(actual = test$TARGET_FLAG, predicted = predict(traditional_cv,newdata=test,type="raw"), prob = 
predict(traditional_cv,newdata=test,type="prob"))
confusionMatrix(eval$predicted, eval$actual, positive = "1")
pred <- prediction(eval$prob.1, eval$actual)
auc <- performance(pred, measure = "auc")@y.values[[1]]
perf <- performance(pred,"tpr","fpr")
plot(perf,main="ROC Curve", sub=paste0("AUC: ",round(auc,3)))
```

@@

```{r, null_prefix = TRUE, comment=NA}
#"KIDSDRIV"         "AGE"             
#[6] "HOMEKIDS"         "YOJ"              "INCOME"           "PARENT1"          "HOME_VAL"        
#[11] "MSTATUS"          "SEX"              "EDUCATION"        "JOB"              "TRAVTIME"        
#[16] "CAR_USE"          "BLUEBOOK"         "TIF"              "CAR_TYPE"         "RED_CAR"         
#[21] "OLDCLAIM"         "CLM_FREQ"         "REVOKED"          "MVR_PTS"          "CAR_AGE"         
#[26] "URBANICITY"       "LOG_INCOME"       "LOG_HOME_VAL"     "AVG_CLAIM"        "PRIOR_ACCIDENT"  
#[31] "COLLEGE_EDUCATED" "URBAN_DRIVER"     "YOUNG_MALE"       "RED_SPORTS_CAR"   "HAS_KIDS"        
#[36] "KID_DRIVERS"     
model <- glm(TARGET_FLAG ~ KID_DRIVERS + MSTATUS + INCOME +
             SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER +ClaimAMTPerYear+ClaimFreqPeryear,
             family = binomial(link = "logit"), over_sample_train)
summary(model)
model_results <- evaluate_model(model, test)

```

```{r}
test$Prediction <- predict.glm(model, newdata=test, type="response")

test$Prediction_Bool <- ifelse(test$Prediction >.5, '1','0')

# Evaluating the model
eval <- data.frame(actual = test$TARGET_FLAG, predicted = test$Prediction_Bool, prob = 
predict(traditional_cv,newdata=test,type="prob"))
confusionMatrix(eval$predicted, eval$actual, positive = "1")
pred <- prediction(eval$prob.1, eval$actual)
auc <- performance(pred, measure = "auc")@y.values[[1]]
perf <- performance(pred,"tpr","fpr")
plot(perf,main="ROC Curve", sub=paste0("AUC: ",round(auc,3)))
```

#Regression

```{r}
model <- lm(TARGET_AMT ~ KID_DRIVERS + MSTATUS + INCOME +
             SEX + CAR_USE + COLLEGE_EDUCATED + REVOKED + URBAN_DRIVER +ClaimAMTPerYear+ClaimFreqPeryear,
             family = poisson, train)

summary(model)
# model_results <- evaluate_model(model, test)

test$Prediction_Amount <- predict(model, newdata=test, type="response")

cases <- test %>%
  complete.cases()
test <- test[cases,]

cor(test$Prediction_Amount,test$TARGET_AMT)

ggplot(test, aes(x=Prediction_Amount, y=TARGET_AMT)) + 
  geom_point()+
  geom_smooth(method=lm)
```
